A load balancing scheme for MMOGs

### 2009-04-08 textos/dissertation/MTAP_LaTeX/pub/tex/latex/svjour3/global/balancing.tex

An initial idea might be to distribute the players between servers, so that the number of players on each server would be proportional to the bandwidth of that server. However, this distribution would not work, for the burden caused by players also depends on how they are interacting with one another. For example, if the avatars of two players are too distant from each other, probably there will be no interaction between them and therefore the server needs only to update each one of them with the result of their own actions. However, if these avatars are close to each other, each player should be updated not only of his own actions, but also of the actions of the other player.

@@

The servers receive the action performed by a player, calculate its outcome and send it to all interested players, who are usually those whose avatars are close to the avatar of the first player. If two players are split between different servers, each of these need not only to send the state update to the player served by it, but it has also to send the update to the server to which the other player is connected. This server, in turn, forwards that state update to the other player (Figure \ref{fig:overheadgeneration}.

@@

Therefore, players who are interacting with each other should, ideally, be connected to the same server. However, it is possible that all players are linked through relations of interaction. For example, two avatars of two different players, may be distant from each other, but both could be interacting with a third avatar, between the other two. However, it is still necessary to divide them between servers. The question is how many pairs of players and which of them will be divided into different server nodes. It is therefore necessary to decide a criterion to group players.

%[TODO: figura da interação de A, B e C]

Next, some principles, used by other authors, will be presented. Based on some of these principles, and on research and implementation carried out, it will be defined the load balancing scheme proposed here.

@@

The main aspect of the solution proposed by the authors was the use of local information (the server that initiated the balancing process and its neighbors), rather than global information (involving all servers in the balance). The former shows small overhead, but it can not solve the problem efficiently in a few steps, as overloaded servers tend to be adjacent. The global approach, in turn, is able to divide the workload in the most balanced way possible, but its complexity grows exponentially with the number of servers involved. The solution is then by involving only a subset of servers, such that its cardinality varies according to the need (if the neighbors of the server which triggered the load balancing are also overburdened, more servers are selected) . Thus, there is a little more information than the local approach, but without the complexity inherent to the global one.

%**	
%To tackle the problem, the authors also subdivide the virtual environment in rectangular \textbf{cells} -- similar to the microcrells. The cells are grouped in \textbf{regions} -- or macrocells -- and each region is managed by one server. Each server keeps its players updated of the state of the virtual environment, while also handling the interactions between avatars in its region. Two cells are adjacent (or neighbors) if they share a border. Similarly, two regions -- and their servers -- are adjacent if there is a pair of adjacent cells, each of which belong to one of the two regions. Figure \ref{fig:leeA} illustrates a distributed server system, consisting of 16 nodes. The virtual environment is divided in 256 cells, which are grouped in 16 regions.

%Para atacar o problema, os autores também subdividem o ambiente virtual em \textbf{células} -- semelhantes às microcélulas -- retangulares, sendo que o número de servidores é muito menor que o número de células. As células são agrupadas em \textbf{regiões} -- ou macrocélulas -- e cada região é gerenciada por um servidor. Cada servidor mantém atualizadas as informações de estado dos usuários e lida com as interações entre avatares na região a ele dedicada. Cada usuário envia e recebe atualizações de estado através do servidor que gerencia a região na qual ele está jogando. Duas células são ditas adjacentes (ou vizinhas) se elas compartilharem uma fronteira. Analogamente, duas regiões, e seus respectivos servidores, são ditos adjacentes se existir um par de células adjacentes, cada uma das quais pertencendo a uma das duas regiões.

To tackle the problem, the authors also subdivided the virtual environment in rectangular \textbf{cells} -- similar to the microcells --, where the number of servers is much smaller than the number of cells. The cells are grouped into \textbf{regions} -- or macrocells -- and each region is managed by one server (Figure \ref{fig:macromicro}. Each server handles the interaction between avatars in the region dedicated to him. It receives the inputs of the players controlling these avatars and sends back to them the up-to-date game state. Two cells are called adjacent (or neighbors) if they share a border. Similarly, two regions, and the servers assigned to them, are called adjacent (neighbors) if there is a pair of adjacent cells, each of which belonging to one of the two regions.

\begin{figure}
 \centering
 \includegraphics[width=0.6\textwidth]{images/macromicro}
 \caption{Cell grouped into four regions (R$_\text{1}$, R$_\text{2}$, R$_\text{3}$ and R$_\text{4}$)}
 \label{fig:macromicro}
\end{figure}

%**
%\begin{figure}
% \centering
% \includegraphics[width=0.8\textwidth]{images/LEE_leeA.jpg}
% \caption{Cells and regions in a distributed server system}
% \label{fig:leeA}
%\end{figure}

%**
The workload of a cell was defined as the number of avatars present in that cell. The authors assumed that all players receive state updates in the same frequency, so that the burden of processing (computing and communication) that a cell requires from a server is proportional to the number of users in that cell. The workload of a region and its designated server is defined as the sum of the individual workloads of the cells which form the region. Each server periodically evaluates its workload and exchange this information with its neighbors. They have assumed, too, that these servers are connected through a high-speed network. Thus, the overhead to exchange workload information among neighbors is limited and considered negligible compared to other costs of the distribution. For the same reason, they also assumed as negligible the overhead of communication between servers in different regions when players are interacting.

@@

A server starts the balancing when the load assigned to it is beyond its power. This server selects a number of other servers to get involved with the distribution. First, it chooses the least loaded server among its neighbors and sends a request that he participates in the load balancing. The chosen server rejects the request if it is already involved in another balancing group, otherwise it responds to the server with the load information of its own neighbors. If the selected neighbor server is unable to absorb the extra workload of the initiating server, the selection is performed again among the neighbors not only of the overloaded server, but also the neighbors of the neighbor chosen in the first step. The selection continues until the workload of the first server can be absorbed -- that is, the workload of all selected servers becomes smaller than a certain limit.

@@

%[TODO:pegar a figura em ingles]

\begin{figure}
 \centering
 \includegraphics[width=1.0\textwidth]{images/LEE_distribution.jpg}
 \caption{Selecting the group of servers for local rebalance}
 \label{fig:lee2}
\end{figure}

@@

%TODO: translate
Há questões de menor importância, mas que também devem ser levadas em consideração e que não foram consideradas em outros trabalhos. Uma delas é que dificilmente o sistema servidor será homogêneo -- considerando-se que é baseado em recursos voluntários. Portanto, não se pode assumir que os servidores tenham a mesma quantidade de recursos. Outro problema, específico do algoritmo proposto por \cite{lee2003sdl}, é que o critério de alocar para cada servidor uma carga equivalente a 90\% da sua capacidade, com o fim de evitar rebalanceamentos constantes, é fraco. Não há garantias de que o sistema servidor terá 11,11\% a mais de capacidade total do que o necessário. Muito pelo contrário, ele deve se adaptar a situações de sobrecarga generalizada. Além disso, o critério de parada é fraco pois o algoritmo não termina se o sistema todo estiver com carga acima de 90\% da sua capacidade, ou se o servidor sobrecarregado não tiver vizinhos disponíveis para balanceamento.

@@

%TODO: translate
Porém, uma idéia importante que foi sugerida é a do uso de grafos para representar o ambiente virtual e de usar algoritmos de particionamento de grafo para realizar o balanceamento de carga. No capítulo \ref{cap:contexto}, foi feita uma introdução a respeito desse princípio, que é a base do esquema de balanceamento de carga aqui proposto. Na seção seguinte, será mostrado como um ambiente virtual de um MMOG pode ser mapeado em um grafo.

%Há questões de menor importância, mas que também devem ser levadas em consideração e que não foram consideradas em outros trabalhos. Uma delas é que dificilmente o sistema servidor será homogêneo -- considerando-se que é baseado em recursos voluntários. Portanto, não se pode assumir que os servidores tenham a mesma quantidade de recursos. Outro problema, específico do algoritmo proposto por \cite{lee2003sdl}, o critério de alocar para cada servidor uma carga equivalente a 90\% da sua capacidade, com o fim de evitar rebalanceamentos constantes, é fraco. Não há garantias de que o sistema servidor terá 11,11\% a mais de capacidade total do que o necessário. Muito pelo contrário, ele deve se adaptar a situações de sobrecarga generalizada. Além disso, o critério de parada é fraco pois o algoritmo não termina se o sistema todo estiver com carga acima de 90\% da sua capacidade, ou se o servidor sobrecarregado não tiver vizinhos disponíveis para balanceamento.

>>>>

An initial idea might be to distribute the players between servers, so that the number of players on each server would be proportional to that server's bandwidth. However, this distribution would not work, for the burden caused by players also depends on how they are interacting with one another. For example, if the avatars of two players are too distant from each other, probably there will be no interaction between them and therefore the server needs only to update each one of them with the result of their own actions. However, if these avatars are close to each other, each player should be updated not only of his own actions, but also of the actions of the other player.

@@

The servers receive the action performed by a player, calculate its outcome and send it to all interested players, who are usually those whose avatars are close to the avatar of the first player. If two players are split between different servers, each of these need not only to send the state update to the player served by it, but it has also to send the update to the server to which the other player is connected. This server, in turn, forwards that state update to the other player (Figure \ref{fig:overheadgeneration}).

@@

Therefore, players who are interacting with each other should, ideally, be connected to the same server. However, it is possible that all players are linked through relations of interaction. For example, two avatars of two different players, may be distant from each other, but both could be interacting with a third avatar, between them. However, it is still necessary to divide the players among servers. The question is how many pairs of players and which of them will be divided into different server nodes. It is therefore necessary to decide a criterion to group players. Many works have been done in the past -- such as \cite{devleeschauwer2005dma,lu2006lbm,chen2005lad,duong2003dls,ahmed2008mol} --  trying to find a near optimal, yet fast, load balancing technique.

In \cite{devleeschauwer2005dma}, for example, a few algorithms are proposed. Their idea is to transfer regions managed by an overloaded servers to  lightly loaded ones, in a way such that the maximum overload among the servers is reduced. Some very interesting principles are used, as the concepts of microcells and macrocells, but the authors' model is weak in some aspects. For instance, it considers the number of players as the load to distribute, heterogeneous systems are not supported, and the algorithm they use to refine an existing partitioning is too general (simulated annealing), considering that there are algorithms specific to partitioning, such as \cite{kernighan1970ehp} or \cite{fiduccia1982lth}, which may perform better. Finally, some of the algorithms proposed in \cite{devleeschauwer2005dma} do not consider the communication between servers.

It is proposed in \cite{lu2006lbm} a load balancing of cpu usage among the nodes of a cluster. The load balancing is done by an ``off the shelf'' NAT load balancer, which distributes the players' connections among the cluster nodes in a round-robin manner. Thus, each cluster node will end up with the same number of players, on the average. This approach, although it does balance the number of clients per cluster node, communication between the server nodes is not considered, for they are connected via a high-speed and low latency network.

Next, some principles, defined in previous works, will be presented. Based on some of these principles, and on research and implementation carried out, it will be defined the load balancing scheme proposed here.

@@

To achieve this goal, the authors also subdivided the virtual environment in rectangular \textbf{cells} -- similar to the microcells --, where the number of servers is much smaller than the number of cells. The cells are grouped into \textbf{regions} -- or macrocells -- and each region is managed by one server (Figure \ref{fig:macromicro}). Each server handles all the interactions between avatars located in the region assigned to it. It receives the inputs of the players controlling these avatars and sends back to them the up-to-date game state.

\begin{figure}
 \centering
 \includegraphics[width=0.4\textwidth]{images/macromicro}
 \caption{Cells grouped into four regions (R$_\text{1}$, R$_\text{2}$, R$_\text{3}$ and R$_\text{4}$)}
 \label{fig:macromicro}
\end{figure}

Two cells are called adjacent (or neighbors) if they share a border. Similarly, two regions -- and the servers assigned to them -- are called adjacent (neighbors) if there is a pair of adjacent cells, each of which belonging to one of the two regions. The workload of a cell was defined as the number of avatars present in that cell. The authors assumed that all players receive state updates of the same lenght and in the same frequency, so that the burden of processing (computing and communication) that a cell requires from a server is proportional to the number of users in that cell. The workload of a region and its designated server is defined as the sum of the individual workloads of the cells which form the region. Each server periodically evaluates its workload and exchange this information with its neighbors. They have assumed, too, that these servers are connected through a high-speed network. Thus, the overhead to exchange workload information among neighbors is limited and considered negligible compared to other costs of the distribution. For the same reason, they also assumed as negligible the overhead of communication between servers in different regions when players are interacting.

The main aspect of the solution proposed by the authors was the use of local information (the server that initiated the balancing process and its neighbors), rather than global information (involving all servers in the balance). When a server is overloaded, it searches for lightly loaded servers close to it in the overlay network. A breadth-first lookup for lightly loaded neighbor servers -- as proposed in \cite{duong2003dls} -- causes a small overhead, but it may not solve the problem efficiently in a few steps, as overloaded servers tend to be adjacent. The global approach, in turn, is able to divide the workload in the most balanced way possible, but its complexity grows exponentially with the number of servers involved.

The solution proposed by the authors is then by involving only a subset of servers, such that its cardinality varies according to the need (if the neighbors of the server which triggered the load balancing are also overburdened, more servers are selected). However, this set is formed in a more clever way than by simply using a breadth-first lookup. The algorithm used to find the servers is described in the next section.

@@

A server starts the balancing when the load assigned to it is beyond its capacity. This server selects a number of other servers to get involved with the distribution. First, it chooses the least loaded server among its neighbors and sends a request that he participates in the load balancing. The chosen server rejects the request if it is already involved in another balancing group, otherwise it responds to the server with the load information of its own neighbors. If the selected neighbor server is unable to absorb the extra workload of the initiating server, the selection is performed again among the neighbors not only of the overloaded server, but also the neighbors of the neighbor chosen in the first step. The selection continues until the workload of the first server can be absorbed -- that is, the workload of all selected servers becomes smaller than a certain limit.

@@

\begin{figure}
 \centering
 \includegraphics[width=1.04\textwidth]{images/localselection}
 \caption{Selecting the group of servers for local rebalance}
 \label{fig:lee2}
\end{figure}

@@

@@
